Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-19 TACAS2018/paper.tex
authors

%-------------------------------
\subsection{State-space representation of LTI models}
\label{sec:ssrepresentation}
%-------------------------------

@@

@@

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs:
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t).
\end{equation}
%
%In this work, 
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in time~\cite{middleton1990digital,van1978computing} 
(cf. Appendix~\ref{sec:cont_aa}) into
\begin{align}
\label{eq:plant}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k, 
\end{align} 
%
where $k \in \mathbb N$ is a discrete counter and $x_{0}=x(0)$ is the initial state. 
Matrices $\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ are equal-sized to $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ respectively,  
and charactirize the discretized plant dynamics. 
%, whereas the latter denote the continuous evolution.  
In this work we make the standard assumption that $\mat{D}_d=\mat{D}=\mat{0}$ \cite{Astrom08}.  
\note{Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.}
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs restricted to $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big( \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big)\;. 
\end{equation}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe the safe state-space.
%
We extend the notation above to convex sets of initial conditions and inputs ($X_0$ and $U$), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x}_0 \in
X_0 \wedge \vec{u} \in U \wedge \vec{x} = \tau(\vec{x}_0,\vec{u})\}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore write $\tau^n(X_0,U)$ to denote the set of states reached from
$X_0$ via $\tau$ in $n$ steps (\emph{$n$-reach set}), i.e.\ for $n\geq 1$
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid &\vec{x}_0 \in X_0 \wedge \forall k\in [0,n-1]\nonumber\\ 
 &\vec{u}_{k} \in U \wedge \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and
vector sums preserve convexity, the sets $X_n = \tau^n(X_0,U)$ are also
convex.

We define the \emph{$n$-reach tube} 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of the reachable sets over $n$ iterations.
%
Moreover, $\hat{X} 
=\bigcup_{n\geq 0} \tau^n(X_0,U)$ 
extends the previous notion over an 
unbounded time horizon.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{State observer}\label{sec:observer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
In most real implementations, the internal state of a plant cannot be
directly measured but rather inferred from its effect on the output by
using an observer.  Essentially, a state observer is a system that
provides an estimate of the internal state of a given real system,
from measurements of the input and output of the real system.

A closed-loop digital control system with observer is shown in Fig.~\ref{fig:observersystem}.
Note that the variables of a state observer are denoted by a "$~\hat{}~$" (e.g. {$\hat {\vec{x}}(k)$})
  to distinguish them from the variables of the actual physical system.
  Moreover, $\vec{y}_e=\hat{\vec{y}_k}-\vec{y}_k$ is the output estimation error defined
  as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, whereas
  $\mat{L}$ is the error sensitivity matrix, which amplifies the output estimation error.
  The figure also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise and will be explained later in the paper.
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.

@@

@@

 Our goal in this work is to synthesise both $\mat{K}$ and $\mat{L}$.

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous time systems}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

 Note that the discretization in Sec.~\ref{sec:ssrepresentation} only 
 allows us to evaluate the state of the system at integer multiples of the sample time.
 However, in order to synthesise $\mat{K}$ and $\mat{L}$, 
 we must be able to ensure that certain properties of the system hold at any given
 time (not just integer multiples of the sample time).
 Consequently, we accelerate the dynamics of the system (according to the model semantics in Sec.~\ref{sec:model_semantics}) in order to compute the state $\vec{x}_T$  at a given time $T$.

>>>>

%-------------------------------
\subsection{State-space representations of LTI models}
\label{sec:ssrepresentation}
%-------------------------------

@@

\note{PLS ELIMINATE BOLD VARIABLES TO DENOTE VECTORS AND MATRICES. BTW, NOTATIONS BELOW ARE NOT CONSISTENT.}

@@

We are interested in control systems with an output $\vec{y}$ that is a linear combination of its states and inputs:
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t).
\end{equation}
%
%In this work, 
Equations~\eqref{eq:ode1}-\eqref{eq:ode2} are soundly discretized in time~\cite{middleton1990digital,van1978computing} \note{mention sample time, referred to later.}
(cf. Appendix~\ref{sec:cont_aa}) into
\begin{align}
\label{eq:plant}
\left\{ 
\begin{array}{l}
\vec{x}_{k+1} = \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} = \mat{C}_d \vec{x}_k + \mat{D}_d \vec{u}_k, 
\end{array}
\right.
\end{align} 
%
where $k \in \mathbb N$ is a discrete counter and $x_{0}=x(0)$ is the initial state. 
Matrices $\mat{A}_d$, $\mat{B}_d$, $\mat{C}_d$ and $\mat{D}_d$ are equal-sized to $\mat{A}$, $\mat{B}$, $\mat{C}$ and $\mat{D}$ respectively,  
and charactirize the discretized plant dynamics. 
%, whereas the latter denote the continuous evolution.  
In this work we make the standard assumption that $\mat{D}_d=\mat{D}=\mat{0}$ \cite{Astrom08}.  
\note{Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.}
%% We synthesize for requirements over this discrete-time domain. 
%% Later, we will address the issue of variable quantization, 
%% as introduced by the ADC/DAC conversion blocks (Fig.~\ref{fig:digitalsystem}).

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, 
starting from an initial set $X_0\subseteq \mathbb{R}^p$, 
with inputs taken from the set $U \subseteq \mathbb{R}^q$, are sequences 
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $, 
%
where $ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = 
\big( \mat{A}\vec{x}_k +
\mat{B}\vec{u}_k \mid \mat{G}\vec{x}_k \leq \vec{h}\big)\;. 
\end{equation}
%
Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs ($X_0$ and $U$, respectively), 
and write $\tau(X_0,U)$ to denote the set of states $\{\vec{x} \mid \vec{x}_0 \in
X_0 \wedge \vec{u} \in U \wedge \vec{x} = \tau(\vec{x}_0,\vec{u})\}$
reached from $X_0$ by $\tau$ in one step. 
%
We furthermore write $\tau^n(X_0,U)$ to denote the set of states reached from
$X_0$ via $\tau$ in $n$ steps (this is also known in literature \note{ref} as the \emph{$n$-reach set}), i.e.\ for $n\geq 1$
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid &\vec{x}_0 \in X_0 \wedge \forall k\in [0,n-1]
%\nonumber\\ &
\vec{u}_{k} \in U \wedge \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}) \} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and vector sums preserve convexity \note{ref}, 
the sets $X_n = \tau^n(X_0,U)$ are also convex.  
We define the \emph{$n$-reach tube} as 
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$ 
generalises the previous notions over an unbounded time horizon.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{State observers}\label{sec:observer}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

In many models of real-world control systems, 
the internal state $x$ cannot be directly measured,  
but is rather inferred from its effect on the output $y$ by using an observer.  
Essentially, a state observer is another dynamical model (below $\hat x$) 
that provides an estimate of the internal state $x$ of the given model, 
from measurements of the input and output of the system. \note{technical details in \ldots}

A closed-loop digital control system with observer is shown in Fig.~\ref{fig:observersystem}. 
Note that the variables of a state observer are denoted by $x_k$. 
  Moreover, $y_e=\hat{y}_k - y_k$ is the output estimation error,  
%  defined as the difference between the real output $\vec{y}_k$ and estimated one $\hat{\vec{y}_k}$, 
  whereas $\mat{L}$ is an error gain matrix, which amplifies the output estimation error. 
Figure \note{\ref{}} also contains references to $\nu_1$, $\nu_2$, $\nu_1'$, $\nu_2'$ and $\nu_t$, which
  denote noise terms that will be explained later.   
%%   With respect to noises (denoted by $\nu$), 
%% $\nu_1$ and $\nu_2$ are the quantization noises introduced by the ADC and the DAC, $\nu_t$ is the noise introduced by the delay, and $\nu_1'$ and $\nu_2'$ are quantization noises caused by roundoffs in the FWL operations.
  Lastly, $k_r$ is a reference gain required by the model.

     Our goal in this work is to synthesise both $\mat{K}$ and $\mat{L}$. \note{K has not been discussed.}

@@

   \& complexnode/.pic={ 
      \node[rectangle,dashed,draw,minimum width=8cm,minimum height=3.5cm,label=\textbf{Model}] (plant)  at ([yshift=-.5cm]dac.center) {};
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (B) at ([xshift=-2.5cm,yshift=.5cm]plant.center)  {\sc $\mat{B}$};
      \node[draw=none] (u) at ([xshift=-1cm,yshift=.15cm]B)  {$u(t)$};
      \node[circle,add={+}{+}{}{}] (p1) at ([xshift=-1.3cm,yshift=.5cm]plant.center) {};
      \node[draw=none] (xdot) at ([xshift=.85cm,yshift=.15cm]p1)  {$\dot{\vec{x}}(t)$};   
      \node[rectangle,draw,minimum width=1cm] (int) at ([xshift=.5cm,yshift=.5cm]plant.center) {\sc $\int$};
      \coordinate (xsouth) at ([xshift=1cm]int);
      \node[draw=none] (x) at ([xshift=1cm,yshift=.15cm]int)  {$\vec{x}(t)$};   
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (C) at ([xshift=2.7cm,yshift=.5cm]plant.center)  {\sc $\mat{C}$};
      \node[draw=none] (y) at ([xshift=1cm,yshift=.15cm]C)  {$\vec{y}(t)$};   
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (A) at ([xshift=.5cm,yshift=-1cm]plant.center)  {\sc $\mat{A}$};
      \coordinate (aeast) at ([xshift=1cm]A);
      \coordinate (awest) at ([xshift=-1.8cm]A);
    }   
    \node[circle,add={+}{+}{}{}]at ([xshift=2cm]C) (lag) {};
    \node[draw=none] (vt)  at ([yshift=-1.5cm]lag) {$\nu_t$};
   complexnode/.pic={ 
      \node[rectangle,dashed,draw,minimum width=8cm,minimum height=5cm,label=\textbf{Observer}] (observer)  at ([yshift=-5cm]plant.center) {};
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (Bd) at ([xshift=-2.5cm]observer.center)  {\sc ${\mat{B}_d}_{\langle C \rangle}$};
      \node[draw=none] (ud) at ([xshift=-1cm,yshift=.15cm]Bd)  {$u_k$};
      \node[circle,add={+}{+}{+}{}] (o1) at ([xshift=-1.3cm]observer.center) {};
      \node[draw=none] (xd) at ([xshift=.85cm,yshift=.15cm]o1)  {$\hat{\vec{x}}_{k+1}$};   
      \node[rectangle,draw,minimum width=1cm] (delay) at ([xshift=.5cm]observer.center) {\sc $z^{-1}$};
      \coordinate (xdsouth) at ([xshift=1.3cm]delay);
      \coordinate (fbsouth) at ([xshift=1.3cm,yshift=-3cm]delay);
      \node[draw=none] (xdp) at ([xshift=1cm,yshift=.15cm]delay)  {$\hat{\vec{x}}_k$};   
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (Cd) at ([xshift=2.7cm]observer.center)  {\sc ${\mat{C}_d}_{\langle C \rangle}$};
      \node[draw=none] (y) at ([xshift=.4cm,yshift=.8cm]Cd)  {$\hat{\vec{y}}_k$};   
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (Ad) at ([xshift=.5cm,yshift=-1.5cm]observer.center)  {\sc ${\mat{A}_d}_{\langle C \rangle}$};
      \coordinate (adeast) at ([xshift=1.3cm]Ad);
      \coordinate (adwest) at ([xshift=-1.8cm]Ad);
      \node[circle,add={-}{}{+}{+}] (o2) at ([xshift=2.7cm,yshift=1.5cm]observer.center) {};
      \node[draw=none] (v1p) at ([yshift=1.2cm]o2) {$\nu_1'$};
      \node[rectangle,draw,minimum width=1cm,minimum height=1cm] (Ld) at ([xshift=.5cm,yshift=1.5cm]observer.center)  {\sc $\mat{L}_{\langle C \rangle}$};
      \coordinate (ldwest) at ([xshift=-1.8cm]Ld);
      \coordinate (bdwest) at ([xshift=-5.5cm]Bd);
      \coordinate (uksouth) at ([xshift=-5.5cm,yshift=5.5cm]Bd);
      \coordinate (o2east) at ([xshift=7.17cm]o2);
    }

@@

 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
 \section{Acceleration of continuous-time models}\label{sec:continuous_time_accel}
 %+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++

The time discretization introduced in Sec.~\ref{sec:ssrepresentation} leads to evaluate the state of the model at integer multiples of the sample time. 
 However, in order to synthesise $\mat{K}$ and $\mat{L}$, 
 we must be able to ensure that certain properties of the system hold at any given
 time (not just integer multiples of the sample time).
 Consequently, we accelerate the dynamics of the system (according to the model semantics in Sec.~\ref{sec:model_semantics}) in order to compute the state $\vec{x}_T$  at a given time $T$.
